---
title: "6. Behavioral Insights from Choice Models"
author: "Heidi Idali Pacheco Almaraz"
format: html
editor: visual
---

```{r}
#| code-fold: true
#| message: false
#| warning: false


library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data
library(gplots) # Various R Programming Tools for Plotting Data
library(ggmosaic) # Mosaic Plots in the 'ggplot2' Framework
library(treemapify)
library(ggridges)
library(ggalluvial)
library(kableExtra) # Construct Complex Table with kable and Pipe Syntax
library(plotly) # Create interactive web graphics
library(tidyr) # Tidy messy data
#library(webshot2) # Take screenshots of web pages

library(evd)
```

1.  What is the difference between a marginal effect and an elasticity?

Un efecto marginal es una medida resumida de la cantidad de cambio en una variable dependiente $y$ cuando una variable independiente $x_k$ cambia en una unidad.

$M_{x_k}^y = \frac{\partial y}{\partial x_k}$

En cambio, una elasticidad resulta útil para resumir la forma en que una variable dependiente $y$ cambia en respuesta a cambios en una variable independiente $x_k$.

$E_{x_k}^y = \frac{\partial y}{\partial x_k}\frac{x_k}{y}$

2.  Why is it inappropriate to calculate the elasticity of a dummy variable?

Porque no tiene cambios marginales definidos, ya que toma valores de 0 o 1.

3.  Use Model 3 in this chapter and calculate the marginal effects and the elasticities for operating cost at the mean of all variables.

```{r}
#| code-fold: true

data("Heating")
glimpse(Heating)

H <- Heating |> mlogit.data(shape = "wide", choice="depvar", varying=c(3:12))

model3 <- mlogit(depvar ~ ic + oc, data = Heating, shape = "wide", choice = "depvar", reflevel = "ec", varying = 3:12)

means <- Heating[1,]

mean_cost <- Heating |> 
  select(starts_with("ic"), starts_with("oc")) |>
  summarise(across(.cols = everything(), mean))

means[3:12] <- mean_cost
```

Efectos Marginales

```{r}
#| code-fold: true

efectos_marginales <- effects(model3, covariate = "oc", type = "ra", data = mlogit.data(means, shape = "wide", choice = "depvar", varying = 3:12))

print("Efectos Marginales:")
print(efectos_marginales)
```

Elasticidades

```{r}
#| code-fold: true
elasticidad <- effects(model3, covariate = "oc", type = "rr", data = mlogit.data(means, shape = "wide", choice = "depvar", varying = 3:12))

print("Elasticidades:")
print(elasticidad)
```

4.  Use Model 3 in this chapter to calculate the rebate needed to reach a 10% penetration rate of heat pumps.

```{r}
penetration_diff <- function(rebate_rate) {
  Heating_mod <- Heating |>
    mutate(ic.hp = ic.hp * (1 - rebate_rate))
  
  heating_mlogit <- mlogit.data(
    Heating_mod, shape = "wide", choice = "depvar", varying = 3:12
  )
  
  pred_prob <- predict(model3, newdata = heating_mlogit)
  mean_hp <- mean(pred_prob[, "hp"])
  mean_hp - 0.10 
}

rebate <- uniroot(penetration_diff, interval = c(0,1), tol = 0.001)
cat("El reembolso necesario para alcanzar una penetración del 10% es del:", rebate$root*100, "%\n")
```

```{r}
penetration_diff <- function(rebate_rate) {
  Heating_mod <- Heating |>
    mutate(ic.hp = ic.hp * (1 - rebate_rate))
  
  prob_pred <- predict(model3, 
                       newdata = mlogit.data(Heating_mod, shape = "wide", choice = "depvar", varying = 3:12), 
                       type = "probabilities")
  
  mean_hp <- mean(prob_pred[, "hp"])
  return(mean_hp - 0.10)
}

rebate <- uniroot(penetration_diff, c(0,0.5))
rebate$root
```

Estimate a new model that extends Model 3 by introducing the age of the household head. Use the electric room system ("er") as the reference level. 

```{r}
Model4 <- mlogit(depvar ~ ic + oc | agehed, Heating, shape = "wide", choice = "depvar", reflevel = "er", varying = c(3:12))

summary(Model4)
```


5. Use the likelihood ratio test to compare your new model to Model 3. Discuss the results. 

```{r}
lrtest(model3, Model4)
```

Con el test podemos observar que la diferencia entre el modelo 3 y el modelo 4 no es muy marcada estadísticamente. Por lo tanto, la edad del jefe del hogar no tiene un impacto significativo en el modelo 3.


6. Is the ratio of the coefficient of installation (or operation) cost to the coefficient of age of household head meaningful. Explain.

Dado que la edad del jefe del hogar no influye estadísticamente en la elección de bombas de calor, la relación entre dicha variable y el costo de instalación u operación tampoco es estadísticamente significativa. 

