---
title: "3. Ejercicio 5. Capitulo 3"
author: "Heidi Idali Pacheco Almaraz"
output: html_document
date: "2025-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Consider the following utility functions for two alternatives, namely i and j:

$$U_i = V_i + \epsilon_i$$
$$U_j = V_j + \epsilon_j$$

Assume that the difference between the error terms below follows the triangle distribution: 

$$\epsilon_q = \epsilon_i - \epsilon_j$$

Parting from the assumption above, derive a binary choice model for the probability of selecting alternative j.


# Función de densidad


```{r}

f_den <- function(x, B=1){
  if(x>=-B & x<=0){
    y <- 1/B^2 *x +1/B
  }else if(x>0 & x<=B){
    y <- -1/B^2 *x +1/B
}else{
    y <- 0
}
  return(y)
}

fv_den <- Vectorize(f_den, "x")

f_dis <- function(x, B=1){
  if(x<(-1)*B){
    y <- 0}else if(x>=-B & x<=0){
    y <- (x+B)*(fv_den(x)/2)
  }else if(x>0 & x<=B){
    y <- 1- (B-x)*fv_den(x)/2
} else{
    y <- 0
}
  return(y)
}


x <- seq(-1.5, 1.5, length.out=500)
y <- sapply(x, f_den)

ggplot()+
  geom_line(aes(x,y), color="purple", , linewidth=1.5)+
  geom_function(fun = fv_den, color="tomato", linewidth=0.5)+
  xlim(-1.5, 1.5)+
  theme_bw()

```


```{r}
decision <- function(Vj, Vi, B=1){
  Vji <- Vj-Vi
  PVj <- f_dis(Vji)
  
  ggplot()+
  geom_function(fun = fv_den, color="tomato", linewidth=0.5, xlim=c(-B-0.1, B+0.1))+
  stat_function(fun = fv_den, geom= "area", fill="green", alpha=0.5, xlim=c(-B, Vji))+ 
  geom_text(aes(B/2, 1/(2*B), label=paste0("Prob(Vj)=", round(PVj,2), sep="")))+
    labs(title="Probabilidad de elección")+
  theme_bw()
  
  
}


decision(Vj=6, Vi=5.5)
```

La probabilidad de seleccionar la alternativa **j** es de **P(Vj) = 0.88** y, por ende, la probabilidad de seleccionar la alternativa **i** es de **P(Vi) = 0.12**

